---
title: "Performance Analytics Plotting Example and Implementation"
author: "Justin Taylor | <taylo5jm@dukes.jmu.edu>"
date: "March 11, 2016"
output: 
  html_document:
    theme: cerulean
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, we walk through a code example in `PerformanceAnalytics` to illustrate how a performance summary chart is currently rendered. Then, we propose an minimal alternate implementation for a plotting function.

# Example
We will walk through a time series plotting example provided by the authors of `PerformanceAnalytics`. This example should create three plots within a graphing window: a wealth index chart, bars for per-period performance, and an underwater chart. The code below has been slightly modified from the original to illustrate my thought process, but the logic remains the same. For example, I wrap `library(PerformanceAnalytics)` in `suppressPackageStartupMessages()` and use a DataTable, instead of `head` to preview the data.  

```{r}
suppressPackageStartupMessages(library(PerformanceAnalytics))
data(managers)
```

```{r}
DT::datatable(managers)
class(managers)
sapply(managers, class)
manager.col = 1
indexes.cols = c(7,8)
```

```{r, fig.width = 9, fig.height = 8, fig.cap = "Wealth index chart, bars for per-period performance, and underwater chart generated in a single plotting window by PerformanceAnalytics"}
charts.PerformanceSummary(managers[,c(manager.col,indexes.cols)], 
                          colorset=rich6equal, lwd=2, ylog=TRUE)
```

Now, we will follow the call stack to see what functions are being used to render this plot. When following the call stack, we will show each function followed by a portion of its definition.  

1. `charts.PerformanceSummary`  

```{r, eval = FALSE}
function (R, Rf = 0, main = NULL, geometric = TRUE,  ...) 
{
    ...
    
    chart.CumReturns(x, main = main, xaxis = FALSE, legend.loc = legend.loc, 
        event.labels = event.labels, ylog = ylog, wealth.index = wealth.index, 
        begin = begin, geometric = geometric, ylab = "Cumulative Return", 
        ...)

    ...
}

<environment: namespace:PerformanceAnalytics>
```

2. `charts.CumReturns`  

```{r, eval = FALSE}
function (R, wealth.index = FALSE, geometric = TRUE, ...) 
{
    ...
  
    chart.TimeSeries(Return.cumulative, colorset = colorset, 
        legend.loc = legend.loc, ...)
}
```

3. `chart.TimeSeries`

```{r, eval = FALSE}
chart.TimeSeries
```

```{r, eval = FALSE}
function (R, auto.grid = TRUE, xaxis = TRUE, yaxis = TRUE, ...) 
{
    ....
  
    for (column in columns:1) {
        lines(1:rows, y[, column], col = colorset[column], lwd = lwd[column], 
            pch = pch[column], lty = lty[column], type = type, 
            ...)
    }
    
    ...
}
<environment: namespace:PerformanceAnalytics>
```

`lines` in Base R is used to render the cumulative returns chart. A similar approach is used in the `timeSeries` package ([https://cran.r-project.org/web/packages/timeSeries/vignettes/timeSeriesPlot.pdf](https://cran.r-project.org/web/packages/timeSeries/vignettes/timeSeriesPlot.pdf))

# Alternate Implementation
In this section, we attempt to reproduce the cumulative returns plot by using `xts::plot.xts` exclusively. We will omit several aspects of the original plot for the sake of brevity, such as adding a legend and fixing axes This document was knit with `RMarkdown`, so the `managers` data set in `PerformanceAnalytics` is already in memory.  

```{r}
library(xts)
library(PerformanceAnalytics)
```

We create a new object, `x`, by subsetting a few columns of `managers`. `HAM1`, `HAM3`, and `HAM4` are now present in `x`. Note, I just picked three arbitrary variables to plot. 

```{r}
x <- managers[,c(1, 3, 4)]
```

Next, we define a function to plot the cumulative returns of each time series. This function is written to deal with the example case in `PerformanceAnalytics` exclusively. A real implementation of this function would have to handle other cases, such as when `x` only inherits from `xts`. Tests would also have to be written with `testthat`. 

```{r}
#' Plot a line graph of cumulative returns for an xts object with one or more variables
#' 
#' @importFrom xts as.xts plot.xts
#' @param x An object that inherits from xts and zoo
#' @return Line graph for a time series object of one or more variables
#' @export plotCumReturns

chartCumReturns <- function(x) {
  stopifnot(inherits(x, c("xts", "zoo")))
  
  x <- xts::as.xts(apply(x, 2, function(y) (cumprod(1 + y))))
  xts::plot.xts(x = x, main = "Cumulative Returns", 
                xlab = "Date", ylab = "Returns")
}

```

Call `chartCumReturns` and pass our `xts` object as an argument.

```{r}
chartCumReturns(x)
```

`lines` still may be called, but `xts::plot.xts` does that under the hood. Once again, this is a very minimal implementation of the original function. The purpose of this exercise was to reproduce the main logic in the `PerformanceAnalytics` function. Obviously, our alternate implementation would have to be generalized for the end user.  

--------------------------------

```{r}
sessionInfo()
```

